On modeling for Kerr black holes: Basis learning, QNM frequencies, and spherical-spheroidal mixing coefficients

### 2018-08-22 docs/mvfits_paper/mvf.tex
misc updates

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
	%
	The ongoing direct detection of gravitational wave signals is aided by representative models of theoretical predictions. In particular, the process of model based detection, and subsequent comparison of signals the general relativity's predictions are aided by the modeling of information related to perturbed Kerr black holes. Here, we summarize recent methods and models for the analytically understood gravitational wave spectra (quasi-normal mode frequencies), and harmonic structure of Kerr black holes (mixing coefficients between spherical and spheroidal harmonics). Towards the construction of these models, two algorithms, GMVP and GMVR, for the automated polynomial and rational modeling of general dimensional complex scalars are presented. GMVR is a quasi-linear numerical method for interpolating rational functions. It therefore represents a solution to the problem of Runge's phenomenon.
	%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% \begin{itemize}
% 	\item BH detections happen, and use fits at various levels
% 	\item Fits for Kerr are important for modeling and testing GR, they are fast to eval comapred to formal calculations which has data analysis implications
% 	\item Many fits have been done for perturbative Kerr parameters: QNM frequencies most prominently.
% 	\item The problem/motivation:
% 		\begin{itemize}
% 			\item 1. There are many fits, and we would like to provide a reference for many of them, including key methods
% 			\item 2. Methods for fits: basis learning
% 			\item 3. (results) Summarize "simple fits": QNM frequencies and decay times, enforcing zero damping
% 			\item 4. (results) Summarize more general fits: mixing coefficients
% 		\end{itemize}
% 	\item (discussion) We have presented ...
% \end{itemize}

@@

In the coming years, expectations for frequent \gw{} detections of increasing \snr{} are high.

@@

Concurrent with \virgo{}, the \aligo{} detectors will enter their third observing run in \check{late 2018}.

@@

At this time, \bbh{} detections are expected at a rate of \check{X} per month.

@@

In this context, signal detection and subsequent inference of physical parameters hinges upon efficient models for source properties and dynamics.

@@

Most prominently, there is ongoing interest in signal models for \bbh{} \imr{}.

@@

As the merger of isolated \bh{s} is expected to result in a perturbed Kerr \bh{}, there is related interest in having computationally efficient models for perturbative parameters, namely those that enable evaluation of the related \rd{} radiation.

@@

These discrete frequencies have associated radial and spatial functions which are \textit{spheroidal} harmonic in nature.

@@

\par In the first and second lines of \eqn{hrd}, we relate the observable \gw{} polarizations, $h_+$ and $h_\times$, with the analytically understood morphology of the time domain waveform.

@@

\par \fig{fig:qnms} displays select training points, as well as model fits for $\cw\lmn$'s real and imaginary parts.

@@

\par In each panel of \fig{fig:qnms}, the horizontal upper and lower demonstrate the effect of mapping $\jf$ onto $\kappa$.

@@

\par Lastly, \fig{fig:qnm_err} shows residual errors.

@@

\par Together, \eqns{eq:cw_fit_1}{eq:cw_fit_9} along with Figures (\ref{fig:qnms})-(\ref{fig:qnm_err}) present precise and accurate fits for the real and imaginary parts of \qnm{} frequencies for gravitational perturbations of Kerr \qnm{s}.

@@

\fig{fig:ys} display comparisons of the fits with training data as well as related residuals.

@@

As with the \qnm{} frequencies, residuals are dominated by small scale oscillations with amplitudes that are fractions of a percent.

%
\section{Discussion}
\label{discuss}

%
\appendix

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  %
  Models of black hole properties are an important component to the ongoing direct detection of gravitational waves from black hole binaries.
  %
  In particular, model based detection, and subsequent parameter estimation are directly enabled by the modeling of information related to perturbed Kerr black holes.
  %
  Here, we summarize recent methods and models for the analytically understood gravitational wave spectra (quasi-normal mode frequencies), and harmonic structure of Kerr black holes (mixing coefficients between spherical and spheroidal harmonics).
  %
  Towards the construction of these models, two algorithms are presented.
  %
  A greedy-multivariate-polynomial (GMVP) regression method and greedy-multivariate-rational (GMVR) regression method are presented for the automated modeling of polynomial and rational functions.
  %
  GMVR is a quasi-linear numerical method for interpolating rational functions. It therefore represents a solution to the problem of Runge's phenomenon.
	%
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

In the coming years, expectations for frequent \gw{} detections of increasing \snr{} are high \cite{TheLIGOScientific:2016pea,Abbott:2016nhf}.

@@

Concurrent with \virgo{}, the \aligo{} detectors will enter their third observing run in approximately late 2018.

@@

During this period, a few to dozens of \bbh{} signals are likely to be detected. \cite{Abbott:2016nhf}.

@@

In this context, signal detection and subsequent inference of physical parameters hinges upon efficient models for source properties and dynamics \cite{Abbott:2016wiq}.

@@

Most prominently, there is ongoing interest in signal models for \bbh{} \imr{} \cite{Blackman:2017dfb, London:2017bcn, Hannam:2013oca}.

@@

As the merger of isolated \bh{s} is expected to result in a perturbed Kerr \bh{}, there is related interest in having computationally efficient models for perturbative parameters, namely those that enable evaluation of the related \rd{} radiation \cite{Berti:2005ys}.

@@

These discrete frequencies have associated radial and spatial functions which are \textit{spheroidal} harmonic in nature \cite{Leaver85,London:2014cma}.

@@

\par In the first and second lines of \eqn{hrd}, we relate the observable \gw{} polarizations, $h_+$ and $h_\times$, to the analytically understood morphology of the time domain waveform.

@@

\par \fig{fig:qnm} displays select training points, as well as model fits for $\cw\lmn$'s real and imaginary parts.

@@

\par \fig{fig:qnm}'s top left and right panels demonstrate the effect of mapping $\jf$ onto $\kappa$.

@@

\par Lastly, \fig{fig:qnm}'s bottom two panels show residual errors.

@@

\par Together, \eqns{eq:cw_fit_1}{eq:cw_fit_9} along with Figures (\ref{fig:qnm})-(\ref{fig:qnm_err}) present precise and accurate fits for the real and imaginary parts of \qnm{} frequencies for gravitational perturbations of Kerr \qnm{s}.

@@

\par \fig{fig:ys} displays fits, training data and related residuals.

@@

As with the \qnm{} frequencies, residuals are dominated by small scale oscillations with amplitudes that are fractions of a percent of the central values.
\section{Discussion}
\label{discuss}
\par We have developed upon previous techniques for the linear and pseudo-linear modeling of low noise data.
In particular, the \gmvp{} algorithm performs multivariate polynomial modeling of real and complex valued scalar functions.
The \gmvr{} algorithm does the same with multivariate rational functions.
\par Treating a toy problem with \gmvr{} demonstrates \gmvr{}'s ability to faithfully recover underlying model parameters for a plausible dataset.
This treatment also demonstrates the convergence of the algorithm's greedy phase with increasing iterations, as well as the convergence of an underlying iterative refinement phase (\ceqn{eq:rat4}).
\par Both \gmvp{} and \gmvr{} may be used to automatically determine the functional form and model for a given dataset that is expected to be respectively multinomial or rational.
An alternative use strategy is to use either \gmvp{} or \gmvr{} to determine a fitting ansatz for individual cases (e.g. individual \qnm{s}), and then use these results to develop a single ansatz for all cases.
This is what as been done for the modeling of \qnm{} frequencies and spherical-spheroidal mixing coefficients.
\par \gmvp{} has been applied to the modeling \qnm{} frequencies. The resulting models have been constrained in the extremal Kerr limit.
These models are found to have residual errors well below $1\%$ when $| \jf | \leq \CwFitCalibrationRegion$.
\appendix
